\Large \bf Dynamic Program Dialysis with Lya

### 2020-01-15 paper/draft.tex

A single application today often incorporates multiple libraries written and published by several different authors.
To make the difficulties of analyzing such library-overreliant more concrete, we describe several example analysis that are often needed by the authors of third-party libraries.
These examples illustrate key requirements for the design of a framework for analyzing applications at the boundaries of modules.
Before describing these examples, however, we offer a brief refresher on the abstractions and corresponding internals of module systems~\sx{bg1}.

@@

\heading{Type Invariant Discovery}
Extracting typing information is helpful in a variety of ways, as type safety is an important property helpful in variety of scenarios. 
%and types are lightweight annotations that can be quickly and efficiently checked.
For example, types can be used to (i) identify program invariants to be preserved during code modifications, (ii) generate (machine-checkable) program documentation, or (ii) guide program synthesis and regeneration~\cite{daikon}.
This is especially problematic in dynamic programming languages:
  in the absence of static type signatures, there is not much one can say about a object \ttt{o} without executing it and introspecting its properties.

As a case in point, consider the \ttt{gRPC} library for serializing and de-serializing objects~\cite{grpc}.
Given a value returned from a library call---say, a bignum or crypto library---a developer has to first run the library manually, take note of the result, and then write a protocol buffer specification file declaring the type; %, common in distributed systems
if the type changes, because of a library version change or because of changes in the .

Dynamic analysis can be used to generate type assertions or invariants.
The static type system is defined as a variant of a simply-typed lambda calculus augmented with union types.
A simple approach would be to define a very simple type system that captures the type of a value by observing the execution of the program---after all, the types are dynamically checked!
The analysis could also capture union types, for cases where it observes that a variable is at times witnessed to hold values of different types---although a single program tends to make calls of the same type signature across its entire lifetime~\cite{daikon}.
The analysis could also do some type ``unpacking'' to deconstruct a type to its structural primitives---for example, noting that an object of type user is comprised of two strings (first and last name) and a number (age).

@@

\subsection{Design Goals}

We have briefly introduced three motivating analyses for extracting information about an application's dependencies.
While the challenges behind these analyses share several characteristics, today they remain largely unaddressed by dynamic analyses frameworks due to a combination of factors:
  these frameworks either require highly specialized solutions often outside the core language,
  do not leverage or offer semantic information at the level of module boundaries, or
  lead to a high (often impractical) performance overheads due to the high granularity of their analyses.
To summarize, these applications would appear to be served well by a system that:
\begin{itemize}
  \item \textbf{(G1)} Operates at the level of modules, leveraging both the semantics and performance potential that this level of granularity potentially offers.
  \item \textbf{(G2)} Does not require learning a new tool or language, but rather allows developers to express their analyses on the same language as the program.
  \item \textbf{(G3)} Can be implemented as a library that bolts onto an existing language runtime environment.
\end{itemize}

The next section describes program dialysis, a new program analysis technique that satisfies these goals, in a language-agnostic manner~\sx{design};
  the section following that describes a concrete implementation for the server-side JavaScript ecosystem~\sx{impl}.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.99\textwidth]{./figs/lya_ex2.pdf}
\vspace{-2mm}
\caption{
  \textbf{Example transformations.}
   Applying runtime transformations (b) and context rebinding (c, d) on a simple \ttt{Math} object~\cf{impl}.
\fixme{No UX!}
}
\label{fig:ex2}
\vspace{-2mm}
\end{figure*}

This section details our implementation of \sys for server-side JavaScript (SSJ) in about 1.5K lines of code on top of version 8.9.4 of the Node.js runtime.
We start with the internals of \sys's Node.js implementation~\sx{impl1}, and then describe the three analyses built on \sys~\sx{impl2}.
We postpone discussion of how other environments diverge from JavaScript for later~\S\ref{reqs}.

In practice, the aforementioned dialysis stages can be implemented in two different ways.
The first is to implement \sys as a modified version of the runtime (\eg \sys-powered JavaScript) in which several stages of its library-loading facility have been augmented in-place with the techniques described earlier.
The second approach is to implement as a third-party library (\eg the \ttt{lya} package) available by the language's package manager.

With both approaches, \sys feels to users as a backward-compatible, drop-in replacement of SSJ's module system---but the second approach has several benefits.
It does not require \sys's users to add a second runtime whose sole purpose is to run analyses;
  it removes \sys's creators and maintainers from the critical path of updates between the runtime developers and its users; 
  and it simplifies \sys's evaluation and comparison with the vanilla runtime (both in terms of performance and correctness).
% However, we have found there are significant opportunities for performance optimization and gains in development complexity by choosing the former approach;
%   we expand on in the discussion section~\sx{reqs}.

@@

Library dialysis, as implemented in \sys, depends on a few features available in the programming language.
In this section we describe how to develop the same solution on other environments.
In the second part of this section, we reflect on some of the current limitations of the system.

\subsection{Applying to Other Environments}
\label{reqs}

\begin{table*}[t]
\center
\footnotesize
\setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Language features used and their availability in other languages}.
    Broadly, the requirements of module dialysis can be broken down into to five elementary requirements.
  }
}
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} lll ll}
\toprule
             &  Instrument \ttiny{import}s  & Traverse objects      & Wrap values         & (Weak) Maps       & Shadow variables      \\
\midrule
Lua          &                              &                       &                     &                   &                       \\
JavaScript   &                              &                       &                     &                   &                       \\
Python       &                              &                       &                     &                   &                       \\
Ruby         &                              &                       &                     &                   &                       \\
Haskell      &                              &                       &                     &                   &                       \\
OCaml        &                              &                       &                     &                   &                       \\
Rust         &                              &                       &                     &                   &                       \\
Java         &                              &                       &                     &                   &                       \\
\bottomrule
\end{tabular*}
\label{tab:compat}
\vspace{-5mm}
\end{table*}

The techniques presented in this paper are applicable to any programming language.
We chose the JavaScript ecosystem to evaluate the feasibility of our approach, primarily because
  (i) it boasts the largest collection of libraries (and number of problems)~\cite{modulecounts} and
	(ii) we had prior experience by writing two systems that make extensive use of dynamic analysis---more sophisticated versions of analyses \fixme{XX} and \fixme{XX}.
In this section, we discuss the requirements for building \sys over other runtime environments.
% FIXME: All these features have to have been discussed earlier

By operating on library boundaries, program dialysis makes use of five key features available in a programming language~\tb{compat}.
All major programming languages can support these five features in a variety of ways, as described below.

First, \sys replaces the implementation of the import statement (\eg \ttt{require}).
% This feature is important for compatibility reasons---otherwise developers would have to replace all import statements in their applications.
If this statement is a function (\eg JavaScript, Scheme, Racket \etc) then replacement can be easily achieved by replacing its definition.
If this statement is a special non-function keyword such as \ttt{import}, this can be achieved by a rewriting rule that replaces import with a special function;
% This is simplified by the fact that the import syntax is trivial.

\sys also needs the ability to interpose on values and transform their components.
This is relatively easy in languages that provide meta-programming capabilities.
Dynamic programming languages offer runtime reflection and conveniently expose object accesses as (over-loadable) functions---\eg \ttt{Proxy} objects in JavaScript, metatables in Lua, metaclasses in Python, and direct-accessor meta-programming Ruby.
Compiled languages such as Haskell, OCaml, and Rust would apply transformations with static metaprogramming facilities, traversing and transforming objects at compile time.
Examples include templates and macro expansions such as Haskell's Template Haskell and Rust's macro system.
And some languages, such as Haskell, Java, and Julia, provide both static and dynamic reflection capabilities, allowing considerable freedom in the choice of implementation.
% Fifth, \sys makes extensive use of interposition which is also available in different programming languages.
% Static programming languages provide different types of support for

\sys makes extensive use of weak maps to support to check for object equality.
Examples of uses include checking for cycles during object traversal and checking whether an object has already been wrapped.
Hash tables that are indexed by reference are supported in virtually all programming languages.
An additional feature useful in garbage-collected languages would be for these (copies of) references to not affect the reclamation of original objects.
Languages either support the ability to create such a weak reference from a normal reference (\eg Python's weak references), or provide structures that automatically hold weak versions of these references (\eg JavaScript's Weak Map).

Finally, \sys makes extensive use of variable shadowing to hide access to the original, unmodified versions of values.
This is a feature that is universally available in programming languages;\footnote{
  A notable exception is the language CoffeeScript. %, which can mimic the same feature using function scope.
}
Deferences in scoping rules---\ie what the shadowed variable ends up referring to upon execution---do not affect the usefulness of this feature, although we have found lexical scoping easier than dynamic scoping.

To conclude, all major languages provide enough facilities to support the dialysis---albeit with a different level of convenience.
Dynamic languages have features---\eg name (re-)binding, value introspection, dynamic code evaluation, and access interposition---that enable runtime transformations.
Runtime transformations conveniently unify module identification with interposition:
  a single function or function-like operator locates a module, interprets it, and applies transformations before exposing its interface in the caller context.
Compiled languages such as Haskell, OCaml, and Rust apply transformations using compile-time metaprogramming facilities.
Compile time transformations alleviate much of the runtime overhead of transformations, improving program startup times;
  and type checking could further aid developers by issuing warnings of incompatible privilege settings.

\subsection{When to Use Dialysis}
\label{principles}
Dynamic analysis is a general analysis technique that discovering facts about programs.
While full-fledged dynamic analysis can provide full visibility into , it may result in higher performance overheads and developer effort.
Dialysis can be thought as a coarser version of analysis stemming from the natural boundaries of libraries, a level of granularity that is just right for certain domains.
Which are these cases exactly?

The first class of  
It is really a quick-and-dirty 
As it quickly isolates homegrown code from standard libraries and third-party code, it is a help pin-point to 
Its availability as a global module makes this trivial.
% This is a particularly good match for dynamic programming languages that are already geared towards quick prototyping.

The second is when the goal is to understand the behavior of third-party code.
Dialysis plugs monitors all around a library, and in clearly delineated abstraction boundaries within that library where the signal-to-noise ratio is naturally expected to be high.
Dialysis deconstructs programs only at the library boundaries, at a granularity well-suited for several tasks.
Examples include the 
N-version programming ---all of which would require a custom tool.
By viewing the program's surrounding environment as libraries---which is the case in all high-level programming languages---dialysis can expand its coverage to the operating system, file system, network and other ``ambient'' services.

Dialysis is \emph{not appropriate} for monolithic programs or ones written in low-level languages such as C.
These limitations is due to several reasons.
First, C programs do not feature clear, ``impenetrable'' boundaries like the ones available in memory- and type-safe languages.
An important feature of the wrapping techniques presented in this paper is that they can ---\ie the full observable behavior around a library.
Unfortunately, this is not true in C where a library can forge a pointer and access arbitrary locations in its the program's address space.
A related problem is the inability to traverse and wrap objects, due to lack of bounds.
It is unclear how to wrap values without any kind of size information.\footnote{
  Recent efforts such as C-Strider~\cite{saur2016cstrider} and ptrSplit~\cite{ptrsplit:17} show promise.
}
% Finally, C is missing direct direct equivalents to \sys's requirements~\sx{reqs}

That said, it is possible to apply dialysis in C programs that have first gone through a process of decomposition or component sanboxing---as is the case in compartmentalization, program slicing, and related areas.
Additionally, many libraries in high-level programming languages today simply wrap C programs either directly or through a foreign-function interface.
Dialysis can be used with these libraries, as it plugs into the abstractions provided by the high-level, wrapper interface.

>>>>

A single application today often incorporates multiple libraries written and published by several different authors.
To make the difficulties of analyzing such library-overreliant software more concrete, we describe three example analyses~\sx{examples}.
These examples illustrate key requirements for the design of a framework for analyzing applications at the boundaries of modules~\sx{goals}.
Before describing these examples, however, we offer a brief refresher on the %abstractions and corresponding
  internals of module systems~\sx{bg1}.

@@

\heading{Type Invariant Discovery}
Extracting type information in a program is helpful in a variety of different ways. %, as type safety is an important property helpful in variety of scenarios. 
%and types are lightweight annotations that can be quickly and efficiently checked.
For example, types can be used to (i) identify program invariants to be preserved during code modifications, (ii) generate (machine-checkable) program documentation, or (iii) guide program synthesis and regeneration~\cite{daikon}.
This is especially important in dynamic programming languages:
  in the absence of static type signatures, there is not much one can say about an object without executing it and inspecting its properties.

As a case in point, consider the \ttt{gRPC} library for serializing and de-serializing objects~\cite{grpc}.
To use it, developers have to provide a protocol-buffer specification describing the types of values that will be serialized.
Given a library---say, bignum or crypto---a developer has to first call it manually, take note of the result's type, and then fill in the protobuf spec. %, common in distributed systems
This process has to be repeated with every change, often due to library updates or changes in the consuming program's structure.

Dynamic analysis can be used to discover such type assertions or invariants.
The analysis would consult the (static) definition of a type \emph{system}, capturing the type of values at the library boundaries by observing the execution of the program.
%% TODO: such as the simply-typed lambda calculus augmented with objects and union types.

@@

\subsection{Design Goals}
\label{goals}

We have briefly introduced three motivating analyses for extracting information about an application's dependencies.
While the challenges behind these analyses share several characteristics, today they remain largely unaddressed by dynamic analyses frameworks due to a combination of factors:
  these frameworks either require highly specialized solutions often outside the core language,
  do not leverage or offer semantic information at the level of library boundaries, or
  lead to a high (often impractical) performance overheads due to the high granularity of their analyses.
To summarize, these applications would appear to be served well by a system that:
\begin{itemize}
  \item \textbf{(G1)} Operates at the level of libraries, leveraging both the semantics and performance potential that this level of granularity can potentially offer.
  \item \textbf{(G2)} Does not require learning a new tool or language, but rather allows developers to express their analyses in the same language as the program.
  \item \textbf{(G3)} Can be implemented as a library that bolts analyses onto an existing language runtime environment.
% \item \textbf{(G4)} It does not incur prohibitive overheads, such that it can be switched on and off selectively during runtime execution.
\end{itemize}

The next section describes program dialysis, a new analysis technique that satisfies these goals, in a language-agnostic manner~\sx{design};
  the section following that describes a concrete implementation for the server-side JavaScript ecosystem~\sx{impl}.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.99\textwidth]{./figs/lya_ex2.pdf}
\vspace{-2mm}
\caption{
  \textbf{Example transformations.}
   Applying runtime transformations (b) and context rebinding (c, d) on a simple \ttt{Math} object~\cf{impl}.
}
\label{fig:ex2}
\vspace{-4mm}
\end{figure*}

% Fixme switch from libraries to modules?

This section details our implementation of \sys for server-side JavaScript (SSJ) in about 1.5K lines of code on top Node.js v8.9.4.
We start with the internals of \sys's Node.js implementation~\sx{impl1}, and then describe the three analyses developed with \sys~\sx{impl2}.
We postpone discussion of how other environments diverge from JavaScript for later~\sx{reqs}.

The aforementioned dialysis design can be implemented in either of two ways.
The first is to implement \sys as a modified version of the runtime (\eg \sys-powered JavaScript) in which several stages of its library-loading facility have been augmented in-place. % with the techniques described earlier.
The second approach is to implement \sys as a third-party library (\eg the \ttt{lya} package) available by the language's package manager.

With both approaches, \sys feels to users as a backward-compatible, drop-in replacement of Node's module system;
  we went with the second approach, as avoiding tight coupling has several benefits:
  it does not require \sys's users to have a second Node copy only for running analyses;
  it removes \sys's developers from the critical path of updates between the Node developers and its users; 
  and it simplifies \sys's comparison with vanilla Node (both in terms of performance and correctness).
The primary drawback was missing a few  opportunities for runtime-performance and development-effort.

@@

Program dialysis depends on a few features provided by the programming language.
In this section, we start by describing its requirements~\sx{reqs} and then reflect on some of its strengths and weaknesses~\sx{principles}.

\subsection{Applying \sys to Other Environments}
\label{reqs}

% \begin{table*}[t]
% \center
% \footnotesize
% \setlength\tabcolsep{3pt}
% \caption{
%   \footnotesize{
%     \textbf{Language features used and their availability in other languages}.
%     Broadly, the requirements of module dialysis can be broken down into to five elementary requirements.
%   }
% }
% \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} lll ll}
% \toprule
%              &  Instrument \ttiny{import}s  & Traverse objects      & Wrap values         & (Weak) Maps       & Shadow variables      \\
% \midrule
% Lua          &                              &                       &                     &                   &                       \\
% JavaScript   &                              &                       &                     &                   &                       \\
% Python       &                              &                       &                     &                   &                       \\
% Ruby         &                              &                       &                     &                   &                       \\
% Haskell      &                              &                       &                     &                   &                       \\
% OCaml        &                              &                       &                     &                   &                       \\
% Rust         &                              &                       &                     &                   &                       \\
% Java         &                              &                       &                     &                   &                       \\
% \bottomrule
% \end{tabular*}
% \label{tab:compat}
% \vspace{-5mm}
% \end{table*}

The techniques presented in this paper are applicable to any programming language.
We chose the JavaScript ecosystem % to evaluate the feasibility of our approach,
primarily because
  (i) it boasts the largest collection of libraries~\cite{modulecounts} and
	(ii) we had prior experience developing more sophisticated version of the security and performance analyses.
We now expand on the requirements for developing dialysis over other runtime environments.
% FIXME: All these features have to have been discussed earlier

By operating on library boundaries, dialysis uses a few special language features, supported by all major programming languages.
First, \sys needs to re-wire the \ttt{import} keyword:
% This feature is important for compatibility reasons---otherwise developers would have to replace all import statements in their applications.
  if it is a function (\eg JavaScript, Scheme, Racket), rewiring amounts to an overwrite;
  if it is a special non-functional keyword, re-wiring needs to be prefixed by a rewriting pass replacing \ttt{import} with a function (at module-loading).
% This is simplified by the fact that the import syntax is trivial.

It also needs the ability to interpose on values and transform their components.
% This is relatively easy in languages that provide meta-programming capabilities.
Dynamic programming languages offer runtime reflection and conveniently expose object accesses as (over-loadable) functions---\eg \ttt{Proxy} objects in JavaScript, metatables in Lua, metaclasses in Python, and direct-accessor meta-programming Ruby.
Compiled languages would apply transformations with static metaprogramming facilities, traversing and transforming objects at compile time.
Examples include templates and macro expansions such as Haskell's Template and Rust's macro system.
% Other languages, such as Java and Julia, provide both static and dynamic reflection capabilities. %, allowing considerable freedom in the choice of implementation.
% Fifth, \sys makes extensive use of interposition which is also available in different programming languages.
% Static programming languages provide different types of support for

\sys makes extensive use of hash-maps---to check for object equality, for cycles during traversal, and for prior wrapping.
Hash-maps are supported by every programming language; many support storing weak references, which avoid affecting object reclamation in garbage-collected environments, and can be created either explicitly (\eg Python) or implicitly by using a ``weak'' structure (\eg JavaScript).
% An additional feature useful in garbage-collected languages would be for these (copies of) references to not affect the reclamation of original objects.
% Languages either support the ability to create such a weak reference from a normal reference (\eg Python's weak references), or provide structures that automatically hold weak versions of these references (\eg JavaScript's Weak Map).

Finally, \sys makes use of variable shadowing to hide the original, unmodified versions of values---a universal feature.\footnote{
  A notable exception is the language CoffeeScript. %, which can mimic the same feature using function scope.
}
% Differences in scoping rules---\ie what the shadowed variable ends up referring to upon execution---do not affect the usefulness of this feature, although we have found lexical scoping easier than dynamic scoping.

To conclude, all major languages provide facilities to support dialysis---albeit with a different level of convenience.
Dynamic languages have features---\eg name (re-)binding, value introspection, dynamic code evaluation, and access interposition---that enable runtime transformations.
Runtime transformations conveniently unify module identification with interposition:
  a single function or function-like operator locates a module, interprets it, and applies transformations before exposing its interface in the caller context.
Compiled languages have other benefits, namely that applying transformations statically ameliorates their runtime overhead.
% and type checking could further aid developers by issuing warnings of incompatible privilege settings.

\subsection{When to Use Dialysis}
\label{principles}
% Dynamic analysis is a general technique for discovering facts about programs.
While heavy-weight dynamic analysis provides full visibility into program execution, it may suffer from higher performance overheads, require significant developer effort, or offer results that are too low-level for the problem at hand.
% Dialysis can be thought as a coarser version of analysis stemming from natural library boundaries---a level of granularity that seems just right for certain problems.
% Which are these cases exactly?

% pressing need, lightweight, 
% multiplies lyas power
Dialysis shines when there is a need for urgency, select online analysis, and a high level of abstraction.
It can help identify (or narrow down) the source of a problem quickly, by swiftly isolating homegrown code from standard and third-party libraries, by requiring a few lines of analysis-specific code inlined, and by leveraging the developer's expertise in their language of choice.
By operating at coarser granularity, using the vanilla optimized runtime, and by toggling parts of the analysis on and off, it provides the ability to perform the analysis online.
Finally, dialysis deconstructs programs only at library boundaries, a naturally coarser granularity that is just the right level for certain problems.
% a granularity that is well-suited for several tasks.
% libraries are a convenient level of abstraction---or the point where abstractions 
% monitors all around a library

% The first class of problems where dialysis shines is quick-and-dirty insights;
% Even if dialysis doesn't solve the problem, it can quickly narrow it down to a certain corner of the application, at which point the cost of a heavy-weight solution is lower.
% % This is a particularly good match for dynamic programming languages that are already geared towards quick prototyping.
% 
% The second is when the goal is to understand the behavior of third-party code.
% Dialysis plugs monitors all around a library, and in clearly delineated abstraction boundaries within that library where the signal-to-noise ratio is naturally expected to be high.
% Dialysis deconstructs programs only at the library boundaries, at a granularity well-suited for several tasks.
% Examples include the 
% N-version programming ---all of which would require a custom tool.
% By viewing the program's surrounding environment as libraries---which is the case in all high-level programming languages---dialysis can expand its coverage to the operating system, file system, network and other ``ambient'' services.

Dialysis is \emph{not appropriate} for monolithic programs or ones written in low-level languages such as C.
These limitations are due to several reasons.
First, C programs do not feature clear, ``impenetrable'' boundaries like the ones available in memory- and type-safe languages.
An important feature of the wrapping techniques presented in this paper is that they can track a library's full observable behavior.
Unfortunately, this is not true in C where a library can forge a pointer and access arbitrary locations in the program's address space.
A related problem is C's inability to traverse and wrap objects;
  due to lack of runtime information about size and structure, it is unclear how to traverse and transform them.\footnote{
  Recent efforts such as C-Strider~\cite{saur2016cstrider} and ptrSplit~\cite{ptrsplit:17} show promise.
}
% Finally, C is missing direct direct equivalents to \sys's requirements~\sx{reqs}

% That said, it is possible to apply dialysis in C programs that have first gone through a process of decomposition or component sanboxing---as is the case in compartmentalization, program slicing, and related areas.
% Additionally, many libraries in high-level programming languages today simply wrap C programs either directly or through a foreign-function interface.
% Dialysis can be used with these libraries, as it plugs into the abstractions provided by the high-level, wrapper interface.
